---
title: "Cell Annotation with TidyComb"
author: "Shuyu Zheng"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cell Annotation with TidyComb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(TidyComb)
```

TidyComb uses information from [ExpasyCellosaurus](https://web.expasy.org/cellosaurus) to generate cell line annotations. As Cellosaurus doesn't provide API, TidyComb embeded whole Cellosaurus data from [Cellosaurus](ftp://ftp.expasy.org/databases/cellosaurus) in the package.

# 1. Check Cellosaurus database version

By defaut, function `CellVersion()` will check the current online Cellosaurus data version

```{r}
CellVersion()
```

Version of the Cellosaurus data embedded by TidyComb can be shown wih:

```{r}
download.file("ftp://ftp.expasy.org/databases/cellosaurus")
CellVersion(file = system.file("cellosaurus.xml", package = "TidyComb"))
```

**Note**: If you find the embedded Cellosaurus data is too old. You can download newest data from ftp://ftp.expasy.org/databases/cellosaurus and save it in your work directory and use this file in the following functions.

# 2. Get cellosaurus accession with cell line name

**Cellosaurus accession** is the unique identifier used by Cellosaurus to archive cell lines

```{r}
cell_name <- c("U87", "HSTS", "LNCAP")
# I'm using the embedded data. You can change is into the file downloaded by yourself
MatchCellAcc(cell_name, file = system.file("cellosaurus.xml", package = "TidyComb")) 
```

# 3. Get cell line information with cellosaurus accession

```{r}
# I'm using the embedded data. You can change is into the file downloaded by yourself
doc <- ParseCell(file = system.file("cellosaurus.xml", package = "TidyComb")) 
GetCellInfo(c("CVCL_0022", "CVCL_0395"), doc)
```

