---
title: "Drug Combination Response Analysis"
author: "Shuyu Zheng"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Drug Combination Response Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(TidyComb)
library(reshape2)
library(dplyr)
```

## 1. Relative Inhibition (RI) for single drug

The **Relative Inhibition (RI)** score is the proportion of the area under the log10-scaled dose-response curve, to the maximum area or the maximum possible inhibition that one drug can achieve at the same dose range.
$$
AUC = \int_{c_1}^{c_2} y_{min} + \frac{y_{max} - y_{min}}{1 + 10 ^ {\lambda(log_{10}{IC_{50} - x'}dx')}} \\
RI = 100 \times \frac{AUC-inh_{min}(c_2 - c_1)}{(1-inh_{min}){c_2 - c_1}}
$$

, where $x' = log_{10}x$; $[c_1,c_2]$ is the concentration range in which the drug was tested; $inh_{min}$ is the minimum inhibition rate the drug can achieve at the dose range [c~1~, c~2~]. (0 is set by default in function)

Input data should be a data frame which contains two columns:

* **dose**: concentration at which drug was tested.
* **response**: percentage inhibition of cell growth at each drug concentrations.

```{r message=FALSE, warning=FALSE}
df <- data.frame(dose = c(0, 0.1954, 0.7812, 3.125, 12.5, 50),
                response = c(2.95, 3.76, 18.13, 28.69, 46.66, 58.82))
sens <- CalculateSens(df)
print(sens)
```

Calculate 95% confidence interval of RI by simulation. The iteration is set as 100 times by default

```{r message=FALSE, warning=FALSE}
pred <- CalculateSens(df, pred = TRUE)$pred
RIConfidenceInterval(pred)
```

# 2. Drug combination analysis

# 1. Function for single drug response matrix

The input data is a matrix which contains the drug combination reaponse value. Its column names are doses of drug added along columns. Its row name are doses of drug added along rows. 

**Note**: the matrix should be sorted by: 
  
1. The concentrations along the column increase _from left to right_;
2. The concentrations along the row increase _from top to bottom_.

```{r}
data <- read.csv(system.file("template.csv", package = "TidyComb"),
                 stringsAsFactors = FALSE)
response.mat <- reshape2::acast(conc_r~conc_c, value.var = "inhibition",
                               data = data[data$block_id == 1, ])
response.mat
```

The `CalculateMat` function with `summary.only = TRUE` setting will return a data frame which contains: 

1. synergy scores calculated with model: ZIP, Bliss, Loewe, HSA
2. Relative IC50 and RI for two drugs added in the combination block.
3. CSS and S score for drug combination sensitivity and synergy, respectively.

```{r message=FALSE, warning=FALSE}
CalculateMat(response.mat, summary.only = TRUE)
```

By setting `summary.only = FALSE`, you can get 3 more tables: 
* **synergy**: percentage inhibitions and synergy scores for each wells in the matrix 
* **surface**: smoothed response or synergy surface for the combination
* **curve**: model used to fit two single drug dose response curves and fitted parameters.

```{r message=FALSE, warning=FALSE}
res <- CalculateMat(response.mat)
res
```

# 2. Function for multiple drug combinations

The input data must be a data frame which contains following columns:

* **block_id**: It is the unique identifier for each drug response (or combination) block.
* **drug_row**: It is the name (or ID) of the drug which was added along the rows of drug combination testing matrix.
* **drug_col**: It is the name (or ID) of the drug which was added along the columns of drug combination testing matrix.
* **cell_line_name**: It is the name (or ID) of cell line on which drug combination response was tested.
* **conc_r_unit**: The unit of concentrations used when adding *drug_row*. For example "uM".
* **conc_c_unit**: The unit of concentrations used when adding *drug_col*. For example "nM".
* **conc_r**: The concentrations of *drug_row* was added in the drug combination test.
* **conc_c**: The concentrations of *drug_col* was added in the drug combination test.
* **inhibition**: The percentage cell growth inhibition at each combinations.

```{r}
data
```

```{r message=FALSE}
CalculateTemplate(data, summary.only = TRUE)
```

# 3. Function for parallel calculation

Function `ParCalculateTemplate` allows you to do the calculation parallelly on multiple cores.

```{r}
res <- ParCalculateTemplate(data, cores = 4)
res
```

