---
title: "Short guidline for TidyComb"
author: "Shuyu Zheng"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{TidyComb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r Loading_package, message=FALSE, warning=FALSE}
library("dplyr")
library("TidyComb")
```

_TidyComb_ is an R package designed for [DrugComb](https://drugcomb.fimm.fi/) data portal. It provides functions to facilitating the processing of drug (combination) dose response data. It involves three main topics: _drug annotation_, _cell line annotation_ and _drug sensitivity analysis_.

This vignette will show a full workstream for processing drug combination data (to make it ready for DrugComb database). Please check other vignettes for more details about _drug annotation_, _cell line annotation_ and _drug sensitivity analysis_.

# Input Data

The input data must be a data frame which contains following columns:

* **block_id**: It is the unique identifier for each drug response (or combination) block.
* **drug_row**: It is the name (or ID) of the drug which was added along the rows of drug combination testing matrix.
* **drug_col**: It is the name (or ID) of the drug which was added along the columns of drug combination testing matrix.
* **cell_line_name**: It is the name (or ID) of cell line on which drug combination response was tested.
* **conc_r_unit**: The unit of concentrations used when adding *drug_row*. For example "uM".
* **conc_c_unit**: The unit of concentrations used when adding *drug_col*. For example "nM".
* **conc_r**: The concentrations of *drug_row* was added in the drug combination test.
* **conc_c**: The concentrations of *drug_col* was added in the drug combination test.
* **inhibition**: The percentage cell growth inhibition at each combinations.

```{r Input_data}
template <- read.csv(system.file("template.csv", package = "TidyComb"),
                     stringsAsFactors = FALSE)
head(template)
```

*Note*: Whithin one block, only one cell_line_name, drug_row, drug_col, conc_r_unit, conc_c_unit. The combination of conc_r or conc_c can not be duplicated in one block. Function *CheckTemplate* was designed for user to check the format of input data. For example:

# Output Data

When uploading new data for DrugComb database, 11 tables are required:
* Manually generate: **study**, **assay**
* Half manually generate: **drug**, **cell_line**, **disease**, **tissue**
* Total automatically generate: **response**, **surface**, **summary**, **curve**, **source**

You can go to the github wiki page ["About_tables"](https://github.com/DrugComb/TidyComb/wiki/About_tables) to find the details about each table.

# 1. Cell line annotation

In this section we will generate tables: 'cell_line', 'tissue', 'disease' and index table 'cell_id' (for facilating table matching in following sections).

## 1.1 Get cellosaurus accession for cell lines.

First of all we need the cell line information from [Cellosaurus](ftp://ftp.expasy.org/databases/cellosaurus). We download the whole data base to work directory and save it as file "cellosaurus.xml".

```{r Download_cellosaurus_data, message = FALSE, eval=FALSE}
# Download "cellosaurus.html" file
download.file("ftp://ftp.expasy.org/databases/cellosaurus/cellosaurus.xml", 
              "cellosaurus.xml")
```

Next we can use downloaded file "cellosaurus.xml" to search the cellosaurus accession for each cell

```{r Match_cell_accession, message=FALSE}
cells <- unique(na.omit(template$cell_line_name)) # The cell name vector must be unique and without NA
cell.match <- MatchCellAcc(cells, file = system.file("cellosaurus.xml", package = "TidyComb"))
print(cell.match)
```

**Note**: Here we have to manually check the matched table and make sure we use the correct cellosaurus accession for further process. Searching conflict cell lines on [ExpasyCellosaurus](https://web.expasy.org/cellosaurus) web page may be helpful.

In this case, cell line _U87_ was matched with two different records. We pick the first one (from "ATCC" company) and generate a "cell line name" - "cellosaurus accession" index table for further data processing.

```{r clean_cell_accession}
cell.match.clean <- cell.match[-2, c("input_name", "cellosaurus_accession")]
```

## 1.2 Generate tables according to cellosaurus accessions.

1. 'cell_line' and 'cell_index' table

Function `GenerateCell` will check cell_line list existing in DruComb database and generate 'cell_line' table and 'cell_line_id' for new cell lines.

```{r Generate_cell_line}
cell <- GenerateCell(cell.match.clean$cellosaurus_accession, 
                     file = system.file("cellosaurus.xml", package = "TidyComb"))
cell$cell_line
cell_index <- cell$cell_id %>% 
  left_join(cell.match.clean, by = "cellosaurus_accession")
```

2. 'tissue' table and 'tissue_index' table for table matching in the following steps.

Function `CheckTissue` will check whether there are new tissues (to DrugComb database) in new 'cell_line' table.

```{r}
new_tisse <- CheckTissue(cell$cell_line$tissue)
```

In this case, the tissue information for cell line "H-STS" is missing. We have to manually include the information into the cell_line table. 

**Note**:

  * the tissue name must be lower cases. 
  * If the tissue name is "central_nevours_system", DrugComb requires more precise classifications: 'brain', 'nervours_system', or 'spinal_cord_or_other_CNS'.

```{r}
cell$cell_line$tissue <- "liver"
new_tissue <- CheckTissue(c(cell$cell_line$tissue))
```

We can see that there is no new tissues. Function `CheckTissue` returns the tissue ID for 'liver' in DrugComb. 

```{r}
tissue_index <- new_tissue$old
tissue_index
```

If there are new tissues detected, function `CheckTissue` will return a list which contains number of total tissues currently in DrugComb, tissues in input data wich already in DrugComb, and new tissues. Whith these information, we can easily generate 'tissue' table for new tissues.

3. 'disease' table and 'disease_index' table for table matching in the following steps.

`CheckDisease` function is same as `CheckTissue` function, but the input data should be a data frame which contains:

* **id**: disease ID
* **dname**: disease name

```{r}
new_disease <- CheckDisease(select(cell$cell_line, id = disease_id, dname = disease_name))
disease_index <- rbind.data.frame(new_disease$old, new_disease$new)
disease_table <- new_disease$new
disease_table
```

4. Cleaning 'cell_line' table

```{r}
cell_line_table <- cell$cell_line %>% 
  left_join(tissue_index, by = c("tissue" = "tname")) %>% 
  rename(tissue_id = "id.y") %>% 
  select(name, synonyms, cellosaurus_accession, disease_id, id = id.x, tissue_id)
cell_line_table
```
  
# 2. Prepare Drug information

## 2.1 Get correct CIDs for drugs

Function `GetCid` using the PubChem API to get the CIDs for input drug names.

```{r Get_CID}
drug_input <- na.omit(unique(c(template$drug_row, template$drug_col)))
drug_cid <- GetCid(drug_input, type = "name")
drug_cid
```

**Note**: Function `GetCid` might return multiple CIDs for some drugs or return `NA` for some drugs. In these cases, user should manually fix the CIDs.

In our case drug "ABT-888" get 2 CIDs. Here we keep the first CID.

```{r}
drug_cid <- drug_cid[!duplicated(drug_cid$input_id), ]
drug_cid
```

## 2.2 Check whether there are new drugs to DrugComb
```{r}
new_drug <- CheckDrug(drug_cid$cid)
```

## 2.3 Get drug information from multiple databases

1. PubChem
```{r Genenrate_drug, message=FALSE}
drug_name <- GetPubNames(cids = new_drug$new)
drug_info <- GetPubchemPro(cids = new_drug$new)
drug <- full_join(drug_name, drug_info)
pub_phase <- GetPubPhase(drug$cid)
drug <- left_join(drug, pub_phase, by = "cid")
```

2. ChEMBL

```{r}
chembl <- GetChembl(drug$inchikey)
drug <- left_join(drug, chembl, by = "inchikey")
```

Merge clinical trail phase from PubChem and ChEMBL. Pick the maximum clinical trail phase.

```{r}
drug$phase <- apply(drug[, c("phase", "chembl_phase")], 1, function(x){
  max(x, na.rm = TRUE)
  })
drug <- select(drug, -chembl_phase)
```

3. UniChem

```{r}
unichem <- GetIds(drug$inchikey)
drug <- left_join(drug, unichem, by = "inchikey") 
drug$chembl_id <- unlist(apply(drug[, c("chembl_id.x", "chembl_id.y")], 1,
                        function(x){
                          if (length(na.omit(x)) == 0){
                            return(NA)
                          } else{
                            return(paste(na.omit(x), collapse = "; "))
                          }
                        }))
drug <- drug[, which(!colnames(drug) %in% c("chembl_id.x", "chembl_id.y"))]
```

4. DrugBank and Stitch

Maching DrugBank and Stitch informations for 'drug' table relys on a local mySQL data base. There is nothing about TidyComb. I'll not show these in the vinegget.

5. Add drug id into drug table
```{r}
drug$id <- seq(new_drug$n + 1, length.out = nrow(drug))
```

6. Arrange the column order of drug table and get the output
```{r}
drug <- select(drug, dname = "name", id, chembl_id, inchikey, smiles, cid, 
               molecular_formula, clinical_phase = "phase", #cid_m, cid_s, stitch_name, 
               drugbank_id = "uni_drugbank", kegg_id = "uni_kegg_c")
drug
```


## 2.4 Generate drug_index table for table matching in the following processes

```{r}
drug_index <- drug_cid %>% 
  left_join(new_drug$old, by = "cid") %>% 
  left_join(select(drug, cid, id), by = "cid")
drug_index$id <- apply(drug_index[, c("id.x", "id.y")], 1, function(x){
  na.omit(x)
})
drug_index <- drug_index[, which(!colnames(drug_index) %in% c("id.x", "id.y"))]
drug_index
```

# 3. calculate synergy scores and generate other files for uploading

```{r Generate_tables, message=FALSE, warning=FALSE}
tables <- CalculateTemplate(template)
names(tables)
```

`pipeline` function generates 5 data frames:
* `response_with_score`: a table containing calculated synergy scores of each interaction block.
* `response`: the **response** table ready for uploading to DrugComb.
* `summary`: the **summary** table ready for uploading to DrugComb.
* `surface`: the **curface** table ready for uploading to DrugComb.
* `curve`: the **curve** table ready for uploading to DrugComb.

# 4. Check and mark dose-response quality for each block

```{r}
bad_quality <- tables$synergy$block_id[which(tables$synergy$inhibition < -200 |
                                               tables$synergy$inhibition > 200)]
bad_quality <- unique(bad_quality)
summary <- tables$summary %>% 
  mutate(quality = rep(NA, n()))
if (length(bad_quality) != 0){
  summary$quality[which(summary$block_id %in% bad_quality)] <- "bad"
}
```

# 5. Replace drug names with drug_id, cell_line_name with cell_line_id

```{r}
curve <- left_join(tables$curve, drug_index, by = c("drug_row" = "input_id")) %>% 
  rename(drug_row_id = "id") %>% 
  left_join(drug_index, by = c("drug_col" = "input_id")) %>% 
  rename(drug_col_id = "id") %>% 
  select(block_id, drug_row_id, drug_col_id, b, c, d, e, model)
curve
summary <- left_join(summary, drug_index, by = c("drug_row" = "input_id")) %>% 
  rename(drug_row_id = "id") %>% 
  left_join(drug_index, by = c("drug_col" = "input_id")) %>% 
  rename(drug_col_id = "id") %>% 
  left_join(cell_index, by = c("cell_line_name" = "input_name")) %>% 
  rename(cell_line_id = "id") %>% 
  select(block_id, drug_row_id, drug_col_id, cell_line_id, conc_r_unit, 
         conc_c_unit, synergy_zip, synergy_loewe, synergy_hsa, synergy_bliss,
         ic50_row, ic50_col, ri_row, ri_col, css_row, css_col, css, S, quality)
summary
```

# Output and check data

```{r}
cell_line_table
tissue_index
disease_table
tables$synergy
summary
tables$surface
curve
```

```{r Output, eval=FALSE, message=FALSE, warning=FALSE}
dir.create("Upload")
write.csv(cell_line_table, "Upload/cell_line.csv", row.names = FALSE)
write.csv(tissue_index, "Upload/tissue.csv", row.names = FALSE)
write.csv(disease_table, "Upload/disease.csv", row.names = FALSE)
write.csv(drug, "Upload/drug.csv", row.names = FALSE)
write.csv(tables$synergy, "Upload/response.csv", row.names = FALSE)
write.csv(summary, "Upload/summary.csv", row.names = FALSE)
write.csv(tables$surface, "Upload/surface.csv", row.names = FALSE)
write.csv(curve, "Upload/curve.csv", row.names = FALSE)
```

Tables "study", "assay" requires manual work for extracting information from data sources. Table "source" relys on these two tables. We can not show the workflow for these 3 tables in this vignette. Please check [github wiki page for TidyComb](https://github.com/DrugComb/TidyComb/wiki) for more details.
