---
title: "Drug Annotation with TydiComb"
author: "Shuyu Zheng"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Drug Annotation with TydiComb}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(TidyComb)
```

# 1. PubChem

## 1. Get CID from PubChem

GetCid queries PubChem database via [PUG REST](http://pubchem.ncbi.nlm.nih.gov/pug_rest/PUG_REST.html) to search matched CIDs of drugs according to other identifiers. Available identifiers are: name, SMILES, InChIKey.

```{r}
GetCid(c("aspirin", "blebbistatin"), type = "name")
GetCid("RDHQFKQIGNGIED-UHFFFAOYSA-N", type = "inchikey")
GetCid("CC1(CCCN1)C2=NC3=C(C=CC=C3N2)C(=O)N.Cl.Cl", type = "smiles")
```

## 2. Get drug name or synonyms 

Drug name and synonyms from CID

```{r}
GetPubNames(c(2244, 1))
```

Drug name and synonyms from Name

```{r}
names <- c("Aspirin", "5-FU")
synonyms <- GetPubSynonymFromName(names)
```


## 3. Get drug property from CID
Drug properties

```{r}
GetPubchemPro(c(2244, 1))
```

Max clinital trial phase: `GetPubPhase

```{r}
GetPubPhase(c(2244, 1))
```

2. ChEMBL

Check the database version information of ChEMBL

```{r}
ChemblVersion()
```

Get ChEMBL ID and max clinical trial with InChI Key

```{r}
GetChembl("PMATZTZNYRCHOR-CGLBZJNRSA-N")
```

3. UniChem

Get DrugBank ID, KEGG compound ID, and ChEMBL ID with InChI Key from UniChem

```{r}
GetIds(inchikey = "BSYNRYMUTXBXSQ-UHFFFAOYSA-N")
```

